### 2022-02-15 paper.tex
in progress, next thing is refs

\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{amsmath,amsfonts,graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary {shapes.geometric}
\usetikzlibrary {positioning}

\begin{document}

\title{Monitoring Time Series With Missing Values:\\a Deep Probabilistic Approach}

\author{Oshri Barazani\inst{1} \and David Tolpin\inst{2}}
\institute{PUB+ \and Ben-Gurion University of the Negev
\\ \email{\{oshribr,david.tolpin\}@gmail.com}}

\maketitle

\begin{abstract}
Systems are commonly monitored for health and security through
collection and streaming of multivariate time series. Advances
in time series forecasting due to adoption of multilayer
recurrent neural network architectures make it possible to
forecast in high-dimensional time series, and identify and
classify novelties early, based on subtle changes in the trends.
However, mainstream approaches to multi-variate time series
predictions do not handle well cases when the ongoing forecast
must include uncertainty, nor they are robust to missing data.
We introduce a new architecture for time series monitoring based
on combination of state-of-the-art methods of forecasting in
high-dimensional time series with full probabilistic handling of
uncertainty. We demonstrate advantage of the architecture for
time series forecasting and novelty detection, in particular
with partially missing data, and empirically evaluate and
compare the architecture to state-of-the-art approaches on a
real-world data set.
\end{abstract}

\section{Introduction}

Modern information systems and operation environments are
commonly monitored through collection and streaming of
multivariate time series. The monitoring tasks comprise both
forecasting, for planning of resource allocation and decision
making, and novelty detection and characterization, for ensuring
faultless functioning and early mitigation of failures and
threats.  Advances in time series forecasting
due to adoption of multilayer recurrent neural network
architectures made it possible to forecast in high-dimensional
time series, and identify and classify novelties (anomalies)
early, based on subtle changes in the trends.  However,
mainstream approaches to multi-variate time series modelling
do not handle well cases when uncertainty is involved, either in
the input, when some of the observations are missing, or in the
output when the distribution of future observations, rather than
their point values, is predicted. For forecast uncertainty
modelling, stochastic latent variable
variants of high-dimensional time series models where
introduced, but so far have had to rely on sampling to account
for uncertainty, limiting the performance of data handling.
Imputation schemes were proposed for dealing with 
missing data, however, they do not generally give a satisfactory
solution in presence of transient unavailability of some of
the data sources (e.g. when a sensor stops working, or a
transport channel malfunctions), which is  a common case with
monitoring of complex systems.

A systematic and theoretically founded approach to handling
both input and output uncertainty would thus constitute a
significant and welcome contribution to the theory and practice
of monitoring of multivariate time series. It would also be
highly desirable for such approach to facilitate efficient
offline (learning) and online (inference) computations. In this
ongoing research, we propose a deep learning architecture which
uses a simple but powerful extension of traditional recurrent
neural network (RNN) architecture which allows both 
\begin{itemize} 
\item to handle missing inputs
in some or all of the components in a multivariate time series,
\item and to accomplish multi-step probabilistic forecasting
\end{itemize}
in high-dimensional time series, paving a path to better
decision making and finer and more robust anomaly detection and
characterization.  We evaluate the architecture on a real-world
data set of multivariate time series collected from a 
network for cloud computing, and empirically demonstrate
advantage of the proposed architecture over commonly used
approaches.

\section{Problem: Multivariate Time Series Forecasting}

The core problem we address is forecasting in a multivariate
time series. Formally, a \textit{time series} is a matrix $X$
of shape $T\times N$, where $T$ is the number of time steps and
$N$ is the number of dimensions. The time steps are assumed to
be equispaced.  A $k$-step probabilistic \textit{forecast}
$\mathcal{F}_{tk}$ at time $t$ is the belief distribution of
time series $X_{t+1:t+k}$ for time steps $t+1 ... t+k$ given the
observed time series $X_{1:t}$ for time steps $1...t$.

The forecasting is accomplished by applying model $\mathcal{M}_\theta$
parameterized by parameters $\theta$ to the observed time
series:
\begin{equation}
\mathcal{F}_{tk} = \mathcal{M}_\theta(X_{1:t})
\end{equation}
The machine learning task is to devise $\theta^*$
that gives the best forecast, in terms of a certain loss
function. A natural loss in the probabilistic setting is the
average negative log likelihood of $\theta$ given a training
data set $\mathcal{X}$ of multiple time series:
\begin{equation}
	\theta^* = \arg\min_\theta \mathbb{E}_{X \in \mathcal{X},t
	\in 1 ... T-k}\left[ -\log \Pr(X_{t+1:t+k}|M_\theta(X_{1:t}))\right]
\end{equation}
When the model is differentiable by $\theta$, the task is
usually accomplished by performing a stochastic gradient loss
minimization.

In the basic case, $X$ is real-valued, $X \in \mathbb{R}^{T
\times N}$. Here, we are interested in an extension of the basic
case, in which some of the elements can be missing from $X$, that is
$X \in (\mathbb{R} \cup \bot)^{T \times N}$.

\section{Architecture: Recurrent Neural Network with Uncertainty Propagation}

\begin{figure}
\begin{subfigure}{.64\textwidth}
\begin{tikzpicture}
\node [rectangle, draw] (rnn) {RNN};

\node (z_t) [above of=rnn] {$z_t$};
\node [trapezium, draw] (mlp) [above of=z_t] {MLP};
\draw [-] (rnn) to (z_t);
\draw [->] (z_t) to (mlp);

%output
\node (musigma) [above of=mlp] {$\mu_{t+1}, \sigma_{t+1}$};
\draw [->] (mlp) to (musigma);

%input
\node (x_t) [below of=rnn] {$x_t$};
\draw [->] (x_t) to (rnn);

\node (h_tm1) [left=12pt of rnn] {$h_{t-1}$};
\draw [->] (h_tm1) to (rnn);

\node (h_t) [right=12pt of rnn] {$h_{t}$};
\draw [-] (rnn) to (h_t);

\node (rnnnext) [rectangle, draw, right=12pt of h_t] {RNN};
\draw [->] (h_t) to (rnnnext);

\node (xnext) [below of=rnnnext] {$x_{t+1}$};
\draw [->] (xnext) to (rnnnext);

%sampling
\node (N) [gray,right=-4pt of xnext] {$\sim \mathcal{N}(\mu_{t+1}, \sigma_{t+1})$};
\draw [->, dashed,gray] (musigma) to [bend left] (N);

\end{tikzpicture}
\caption{Conventional model}
\label{fig:dt-conventional-model}
\end{subfigure}%
\begin{subfigure}{.36\textwidth}
\begin{tikzpicture}
\node [rectangle, draw] (rnn) {RNN};

\node (z_t) [above of=rnn] {$z_t$};
\node [trapezium, draw] (mlp) [above of=z_t] {MLP};
\draw [-] (rnn) to (z_t);
\draw [->] (z_t) to (mlp);

%output
\node (musigma) [above of=mlp] {$\mu_{t+1}, \sigma_{t+1}$};
\draw [->] (mlp) to (musigma);

%input
\node (x_t) [below of=rnn] {$\mu_t=x_t,\sigma_t=0$};
\draw [->] (x_t) to (rnn);

\node (h_tm1) [left=12pt of rnn] {$h_{t-1}$};
\draw [->] (h_tm1) to (rnn);

\node (h_t) [right=12pt of rnn] {$h_{t}$};
\draw [-] (rnn) to (h_t);

\node (rnnnext) [rectangle, draw, right=12pt of h_t] {RNN};
\draw [->] (h_t) to (rnnnext);

\node (xnext) [below of=rnnnext] {$\mu_{t+1}, \sigma_{t+1}$};
\draw [->] (xnext) to (rnnnext);

\node (F) [right=12pt of z_t] {};
\draw [-,dashed,gray] (musigma) to [bend left] (F);
\draw [->,dashed,gray] (F) to [bend right] (xnext);

\end{tikzpicture}
\caption{Model with uncertainty propagation}
\label{fig:dt-uncertainty-propagation}
\end{subfigure}
\caption{Time series models}
\label{fig:dt-time-series-models}
\end{figure}

A popular realization of the forecasting model
$\mathcal{M}_\theta$ is a recurrent neural network (RNN), with
$\theta$ corresponding to the network parameters.  There is a
range of neural recurrent models of varying complexity to deal
with time series forecasting. Most models include a recurrent
unit which threads the state through the time steps, accepts
data as inputs and produces next step predictions as outputs.
The simplest model is an RNN with a fully-connected readout
layer to produce forecasts
(Figure~\ref{fig:dt-conventional-model}).  RNN can be based on
LSTM~\cite{}, GRU~\cite{}, or another architectural variant, and
is often multi-layer.  Architectures  may also include
intermediate modules and sampling-based variational
layers~\cite{}. The overall architecture stays almost the same,
with more connections, intermediate modules and sampling-based
variational layers.

\paragraph{Input and Output} This architecture normally accepts
data vectors and outputs vectors of predictions of means and
standard deviations (the output is twice as wide as the input).
Then, to generate a forecast, one samples from
$$\mathcal{N}(\mu, \sigma)$$ for each component, and feeds the
sample as the input to the next time step.

\paragraph{Training} The model is trained to maximize
probability of prediction. In the most basic case, a single step
is predicted for each time step in the series. If a series has
10 steps, then steps 1 .. 9 are the input, and the steps 2 .. 10
are the truth. The network is trained to minimize negative log
probability of true observations given predictions under
variational approximation of product of Gaussians.

\paragraph{Novelty detection} After training, the model is used
to produce predictions of future observations. There are two
related but different phenomena indication a novelty (anomaly)
in time series behavior:
\begin{enumerate}
\item Predicted volatility of the time series is high, that is, future
observations can only be forecast uncertainly (with high variance). 
\item Probability of actual observations, when observed, given a prediction
from a past state, is low.
\end{enumerate}
Either phenomenon, or both of them, can be used to alert about
novelties in the time series. In recurrent neural network
architectures, the hidden state ($h_t$ in
Figure~\ref{fig:dt-time-series-models}) can be used to identify
and classify anomalies.

However, the basic scheme outlined above poses difficulty in
applications with high-dimensional time series and partially
missing observations. Sampling based uncertainty assertion
impacts performance, and missing observations are often imputed
in \textit{ad hoc} manner~\cite{}. An architecture which
incorporates confidence about data and in which observed and
predicted data are interchangeable is  highly desirable. For
example, if out of 5 components 3 were measured and 2 predicted
from an earlier step we want to input them both into the next
time step for further forecasting. In addition, the model
architecture should be capable of robust uncertainty prediction
and benefit from training with multiple steps of out-of-sample
data.

To achieve these objectives, \textbf{we propose the following
changes} (Figure~\ref{fig:dt-uncertainty-propagation}):

\begin{enumerate}
\item The input, as well as the output, is a vector of means followed
by standard deviations. If the data has 5 components, the input
will be 10-dimensional. For observed data --- measurements
present at the current time step --- the standard deviation is
zero. For missing data the input is the mean and the standard deviation
as predicted from the preceding time steps.

\item Since we introduce confidence into the input, we cannot
train the network myopically, on a single step prediction --- or
the standard deviations in the input data will always be zero,
and the network will never learn how to use them. To overcome
this, we train on multiple predicted steps. We feed each
prediction, without sampling, as input to the next step and
compute the loss as negative log probability of this number of
future points versus our prediction.
\end{enumerate}

To illustrate, given the data set of 5 dimensions, the input has 10
dimensions. If we train with 3 time steps lookahead, the ground
truth will be a matrix of size $3\times 5$. The prediction against
which the likelihood of this ground truth is computed will be
a matrix of size $3 \times 10$. Intuitively, we would expect the
predicted standard deviation to increase along the time axis
for each component.

The ability of probabilistic forecasting with uncertainty, in
the form of a multivariate Gaussian distributions, far into the
future, opens opportunity for application for more robust
novelty detection approaches. Instead of detecting novelty based
on log probability of observations given predictions from the
past, which is prone to false positives due to observation
noise, novelties can be detected and analysed by comparing
predictions of the same time point from different points in the
past. In this case, KL-divergence between predictions provide a
theoretically sound and robust mechanism for detection of
anomalies, and is in particular relevant for monitoring of large
operation environments with high dimensionality of time series
and occasional missing values and heteroskedastic noise.

\section{Case Study: Monitoring a Computer Cloud}

\section{Related Work}

\section{Discussion and Future Research}

\section*{Acknowledgements}

\bibliographystyle{splncs04}
\bibliography{refs.bib}

\end{document}